### 2011-07-20 Text/paper.tex
Need to resolve the two-dimensional cap contributions by automodel solutions.

\section{Mass Transfer Correlations}
We examine the following three characteristic correlations:
\begin{description}
\item[I]
In what follows we want to address the following empirical correlation for the volumetric mass
transfer coefficient addressed in the recent
publication by \citet{yue-mass}:
\begin{equation}
k_L a=\frac{2}{d_h} \Bigl(\frac{D
U_{\mathrm{bubble}}}{L_{\mathrm{bubble}}+L_{\mathrm{slug}}}\Bigr)^{0.5}
\Bigl(\frac{L_{\mathrm{bubble}}}{L_{\mathrm{bubble}}+L_{\mathrm{slug}}}\Bigr)^{0.3}.
\end{equation}
This correlation is within $10\%$ accuracy for experimental results within the following range of
parameters: bubble velocity $0.4\,\mathrm{m/s}<U_{\mathrm{bubble}}<2\,\mathrm{m/s}$, bubble length
$1.4<\frac{L_{\mathrm{bubble}}}{d_h}<6.3$, slug length $1<\frac{L_{\mathrm{slug}}}{d_h}<3.2$,
hydrodynamic diameter $d_h = 400
\mathrm{\mu m}$.

For long slug lengths in circular capillaries, where the mass transfer from bubble caps is
predominant, the folowing correlation is derived by
\citet{bercic-mass}:
\begin{equation}
k_L a = \frac{0.111
(U_{\mathrm{gas}}+U_{\mathrm{liq}})^{1.19}}{\bigl((1-\varepsilon_{\mathrm{gas}})(L_{\mathrm{bubble}}
+L_ {\mathrm{slug}} )\bigr)^{0.57} },
\end{equation}
where $\varepsilon_{\mathrm{gas}}=\frac{U_{\mathrm{gas}}}{U_{\mathrm{bubble}}}$ is the gas holdup,
$U_{\mathrm{gas}}$ is the superficial gas velocity, and $U_{\mathrm{liq}}$ is the superficial
liquid velocity. The superficial liquid velocity is calculated with the steady state condition in
the middle of the slug as:
\begin{equation}
U_{\mathrm{liq}}=\frac{\oint U \mathrm{d}A}{A}.
\end{equation}
The superficial liquid velocity can be calculated through the calculated bubble volume and gas
holdup:
\begin{equation}
\begin{aligned}
&\varepsilon_{\mathrm{gas}}=\frac{\text{bubble area}}{15 H^2}\\
&U_{\mathrm{gas}}=\varepsilon_{\mathrm{gas}} U_{\mathrm{bubble}}.
\end{aligned}
\end{equation}

\item[III]
\citet{vanbaten-circular} performed numerical simulations for circular capillaries by assuming that
there is a separate contribution of the liquid film and the bubble caps to mass transfer. Their
correlation is as:
\begin{equation}
k_L a = \frac{2}{\sqrt{\pi}}\sqrt{\frac{D U_{\mathrm{bubble}}}{(L_{\mathrm{bubble}}-d_h)}}
\frac{4(L_{\mathrm{bubble}}-d_h)}{d_h(L_{\mathrm{bubble}}+L_{\mathrm{slug}})}+2\frac{\sqrt{2}}{\pi}
 \sqrt{\frac{D U_{\mathrm{bubble}}}{d_h}} \frac{4}{(L_{\mathrm{bubble}}+L_{\mathrm{slug}})}.
\end{equation}
This correlation is valid for short contact times, when the mass transfer is important from bubble
caps as well as from the film. Short contact times are defined through Fourier number as:
\begin{equation}
Fo=\frac{D t_{\mathrm{film}}}{\delta^2},
\end{equation}
where $t_{\mathrm{film}}=\frac{L_{\mathrm{film}}}{U_{\mathrm{bubble}}}$ is the time exposure of the
bubble while it propagates through distance $L_{\mathrm{film}}$.

\end{description}
Our goal is to compare numerical simulations with given correlations.

\section{Mass transfer}
The mass transfer is characterized by non-dimensional numbers as Schmidt number (viscous diffusion
rate over the mass diffusion rate), the Peclet number (convection over diffusion) and the Sherwood
number (convection mass transfer over the diffusion mass transfer):
\begin{equation}
\begin{aligned}
&Pe=\frac{U L}{D}\\
&Sh=\frac{K L}{D}\\
&Sc=\frac{\nu}{D}\\
\end{aligned}
\end{equation}

@@

In what follows we will perform the matching of obtained hydrodynamics results
\cite{kuzmin-binary2d} with physical ones from experiment of \citet{bercic-mass}. The diffusion
coefficient of the methane under standard conditions \cite{methane-properties} is
$1.84\times 10^{-5} \mathrm{cm}^2/\mathrm{s}$. The diameters of capillaries were as $1.5$, $2.5$ and
$3.1$ $\mathrm{mm}$. For simplicity reasons we will choose the diameter of capillary as $1.5\,
\mathrm{mm}$. 
We will match parameters through given non-dimensional numbers. From the hydrodynamics
simulations the following parameters supplied:
\begin{equation}
\begin{aligned}
Ca=\frac{U \mu_{\mathrm{liq}}}{\gamma}\\
Re=\frac{\rho_{\mathrm{liq}} U L}{\mu_{\mathrm{liq}}}.
\end{aligned}
\end{equation}
From these two equation one can obtain the physical velocity obtained from non-dimensional numbers:
\begin{equation}
\begin{aligned}
Ca\cdot Re= \frac{\rho_{\mathrm{liq}} U^2 L}{\gamma}\\
U=\sqrt{\frac{Ca\,Re\,\gamma}{\rho_{\mathrm{liq}}L}}
\end{aligned}
\end{equation}

Assuming that the surface tension of water is $0.0728\,\mathrm{N/m}$, the density is
$1000\,\mathrm{kg/m^3}$,the hydraulic diameter of the channel to be $1.5\,\mathrm{mm}$ one can
calculate the physical velocity of the bubble.
Table \ref{table:twod:simulations} represents the number of simulations runs for the grid as
$202\times 3002$ and its nondimensional parameters. One can see that velocity is quite reasonable
as \citet{bercic-mass} declared that experiments were conducted for velocities in the ranges of
$0.01$ to $0.4\,\mathrm{m}/\mathrm{s}$ but for less capillary numbers. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
$Ca$    &$Re$     &$U_{LB}$ &$\delta$&$\varepsilon_{\mathrm{gas}}$
&$U_{\mathrm{phys}}\mathrm{m/s}$\\
$0.026$ &$0.449$  &$0.0014$ &$0.040$ &$0.306$ &$0.023$   \\ 
$0.047$ &$0.820$  &$0.0027$ &$0.058$ &$0.293$ &$0.043$   \\ 
$0.080$ &$1.378$  &$0.0045$ &$0.085$ &$0.280$ &$0.073$ \\
$0.065$ &$1.126$  &$0.0037$ &$0.076$ &$0.266$ &$0.059$      \\
$0.222$ &$3.807$  &$0.0125$ &$0.122$ &$0.253$ &$0.202$  \\
$0.479$ &$8.222$  &$0.0271$ &$0.151$ &$0.249$ &$0.437$  \\
$0.736$ &$12.617$ &$0.0416$ &$0.164$ &$0.236$ &$0.671$  \\ 
$0.989$ &$16.960$ &$0.0559$ &$0.172$ &$0.230$ &$0.902$  \\
\end{tabularx}
\caption{Two-dimensional simulations for different
capillary numbers. \label{table:twod:simulations}}
\end{table}

\section{Lattice Boltzmann implementation}
For the lattice Boltzmann implementation we took $\omega=1.99$ which results in the following
lattice Boltzmann diffusivity parameter:
\begin{equation}
\begin{aligned}
&D=\frac{1}{3}\Bigl(\frac{1}{\omega}-\frac{1}{2}\Bigr)\\
&D=0.0008375
\end{aligned}
\end{equation}
In terms of Schmidt number (taking the liquid viscosity in former simulations as $\frac{2}{3}$) that
can result in:
\begin{equation}
\begin{aligned}
&Sc=\frac{\nu_{\mathrm{liq}}}{D}\\
&Sc=796\\
\end{aligned}
\end{equation}
Taking the water as $8.9 \times 10^{-4}\,\mathrm{Pa\cdot s}$ or kinematic viscosity as $8.9 \times
10^{-7} \,\mathrm{m^2/s}$ and the Schmidt number as $796$ that results in the following quantity:
\begin{equation}
D=1.118\times 10^{-5}\,\mathrm{cm^2/s},
\end{equation}
which is less than the diffusivity of the methane $1.84\times 10^{-5}\,\mathrm{cm^2/s}$ but is
physical. After calculation of physical parameters one can calculate characteristics for the mass
theoretical mass transfer. Table \ref{table:lengths} gives nondimensional and physical lengths of a
bubble and a slug.
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$&Bubble length& Slug Length&
$U_{\mathrm{gas}}\,\mathrm{m/s}$&$U_{\mathrm{liq}}\,\mathrm{m/s}$&$a,\,\mathrm{m^{-1}}$\\
\hline
$0.026$&$5.225$&$9.775$ &$0.0070$ &$0.0209$&$2.5538$\\
$0.047$&$5.22$ &$9.78$  &$0.0126$ &$0.0379$&$2.5431$\\
$0.080$&$5.215$&$9.785$ &$0.0204$ &$0.0615$&$2.5326$\\
$0.065$&$4.965$&$10.035$&$0.0157$ &$0.0504$&$2.4142$\\
$0.222$&$5.25$ &$9.75$  &$0.0511$ &$0.1560$&$2.5253$\\
$0.479$&$5.565$&$9.435$ &$0.1089$ &$0.3155$&$2.6518$\\
$0.736$&$5.51$ &$9.49$  &$0.1584$ &$0.4677$&$2.6269$\\
$0.989$&$5.505$&$9.495$ &$0.2074$ &$0.6190$&$2.6340$\\
\hline
\end{tabularx}

\caption{The nondimensional (scaled to the height of the channel) lengths for the same
lattice Boltzmann simulations as in Table \ref{table:twod:simulations}. As well superficial
gas-liquid velocities are presented.\label{table:lengths}}
\end{table}

The literature correlations of mass transfer dependence on the bubble velocities are presented in
Fig. \ref{fig:mass:transfer:theoretical}. Note, that the correlations presented in
Fig.\ref{fig:mass:transfer:theoretical} are used primarily for circular capillaries
(\citeauthor{vanbaten-circular,bercic-mass}) and for square capillaries (\citeauthor{yue-mass}). We
adapt these results for flow between parallel plats by using hydraulic
diameter \cite{bercic-mass} which can be calculated as:
\begin{equation}
d_h = \lim_{W\to\infty}\frac{4A}{P}=\lim_{W\to\infty}\frac{4 W\,H}{2(W+H)}=2 H.
\end{equation}
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/theoretical_correlations.eps}
\caption{The literature correlations by various authors. One
can see the large deviation among them.\label{fig:mass:transfer:theoretical}}
\end{figure}

The volumetric mass transfer coefficient is calculated according to the work of
\citet{vanbaten-circular}. The concentration flux is calculated as the difference between overall
average concentration taken in the whole domain ($C_{overall}=\int_{V} C \mathrm{d}V /V$) 
{\color{red} (In this particular moment I don't know whether to take the Volume of whole domain or
just the liquid phase. I will go for whole domain.)}
at time
$t_1$ and time $t_2$ divided on the time difference $t_2-t_1$. Then the volumetric mass transfer
coefficient is calculated as:
\begin{equation}
\label{main:simulation:equation}
k_L a=\frac{\text{Flux}}{C_s-C_{outlet}} \frac{\text{bubble surface area}}{\text{unit cell volume}},
\end{equation}
where $C_{outlet}=\int{C U_{outlet}\mathrm{d}A}/\int{U_{outlet}\mathrm{d}A}$. 
{\color{red} Do not understand why we need this ratio bubble surface area/unit cell volume. In this
case the dimensional units do not coincide. In the calculations I will avoid it until further
clearance.}
To properly do
non-dimensioanalization one needs to insure the time conversion factor when the flux is calculated
inside the lattice Botlzmann system. The time conversion can be easily obtained:
\begin{equation}
\begin{aligned}
&U_{LB}=U_{phys}\frac{\Delta t}{\Delta x}\\
&\Delta t=\frac{U_{LB}}{U_{phys}}\Delta x\\
&\Delta t=\frac{1.5\,\mathrm{mm}}{200} \frac{0.0559}{0.902\,\mathrm{m/s}}=4.64\times
10^{-7}\,\mathrm{s} 
\end{aligned}
\end{equation}
We as well need to address the concentration being in a whole system non-dimensionalized towards the equilibrium:
if the domain size is

The bubble surface length is the perimeter of the bubble in two-dimensional space and it is
calculated with the extraction of the bubble countour and representing it with the polygon. After
that the calculated perimeter is the effective summation of line lengths and ca be represented as:
\begin{equation}
P=\sum_i{\sqrt{(x2_i-x1_i)^2+(y2_i-y1_i)^2}},
\end{equation}
where the index $i$ represents the polyline consisting of two points $(x1,y1)$ and $(x2,y2)$.

The corresponding volumetric mass transfer coefficients depending on the velocity are indicated in
Table \ref{table:volumetric:coefficients}.

\section{Simulation results}
\subsection{Steady-state}
Using Eq.\ref{main:simulation:equation} we calculated the mass transfer coefficient. One can see
the plots for the mass transfer coefficient versus time in Fig. \ref{fig:steady:state}. We state as
a steady-state condition the average of concentration values between $500000$ to $1000000$ time steps. The values are presented 
\begin{figure}
\includegraphics[width=\textwidth]{Figures/steady_state.eps}
\caption{The mass transfer coefficient vs. time iterations. One can see that the pseudo steady
state can be reached after $500000$. The mass transfer coefficient is taken as the average of concentration values between $500000$ to $1000000$ time steps..\label{fig:steady:state}}
\end{figure}

\section{To revise:}
Then the diffusion penetration length is:
\begin{equation}
\delta_{penetration}=\sqrt{D t_{film}}=\sqrt{\frac{D L_{bubble}}{U_{bubble}}}.
\end{equation}
The bubble length is $5$ of diameters. Thus, the bubble length is $7.5\,\mathrm{mm}$. Therefore:
\begin{equation}
\delta_{penetration}=\sqrt{\frac{1.84 10^{-9} \mathrm{m^2/s} 7.5 10^{-3} \mathrm{m}}{4.3 \time
10^{-2}\mathrm{m/s}}}=1.8 \times 10^{-5}
\end{equation}
With the given value of $200$ nodes to resolve the diameter for the given simulation, that results
in $2.4$ nodes, which is small but considerable number to conduct the simulation.
Foirier number which is defined as:
\begin{equation}
Fo=\frac{D t_{film}}{\delta_{film}^2}=\frac{1.84\times 10^{-9} 0.1744}{\times
(8.7\times10^{-5})^2}=0.0423
\end{equation}
As it is indicated by \citet{vanbaten-circular} the Fourier number $Fo<0.1$ stays for short
contact, the film is unsaturated. 
Thus, one can calculate the corresponding parameters:
\begin{equation}
Pe_{min}=\frac{10^{-2} 1.5 10^{-3}}{1.84  10^{-9}}=0.81 \times 10^{4}\gg 1
Sc_{min}=\frac{10^{-6}}{1.84 10^{-9}}=543,
\end{equation}
which imply the Reynolds number to be as:
\begin{equation}
Pe=Re Sc,\,Re=Pe/Sc=14.91
\end{equation}

The Schmidt number $Sc \approx 500$ states for dye in the liquid and assumes the dominance of
convection over diffusion. In terms of the lattice Boltzmann system, the liquid viscosity equals to
$\frac{1}{3}(\tau_{liq}-\frac{1}{2})=\frac{2}{3}$. That means that the relaxation rate for the
advection-diffusion equation should be as:
\begin{equation}
\frac{1}{3}\Bigl(\tau_{\phi}-\frac{1}{2}\Bigr)=\frac{\nu}{Sc}\\
\tau_{\phi}=0.50368,
\end{equation}
which is hard to simulate with the lattice Boltzmann equation, but possible . As it was discussed
\cite{giavedoni-numerical} the inertia can be neglected for the Reynolds number $Re<70$ for bubble
configuration, but not for the hydrodynamics simulations.

The easy way to do is not to match the Peclet number $Pe$ along with the Reynolds number, leaving
only matching the Schmidt number. However, the thorough study needs to be conducted to perform the
matching parameters as $Ca$, $Re$, $Sc$ and $Pe$ numbers.

If we match only the Peclet number then the Reynolds number can be changed as to adjust the
velocity multiplied on coefficient, keeping Reynolds number small and anticipating results to see.

>>>>

\section{Mass Transfer Correlations}
\label{sec:correlations}
We examine the following three characteristic correlations:
\begin{description}
\item[I]
In what follows we want to address the following empirical correlation for the volumetric mass
transfer coefficient addressed in the recent
publication by \citet{yue-mass}:
\begin{equation}
k_L a=\frac{2}{d_h} \Bigl(\frac{D
U_{\mathrm{bubble}}}{L_{\mathrm{bubble}}+L_{\mathrm{slug}}}\Bigr)^{0.5}
\Bigl(\frac{L_{\mathrm{bubble}}}{L_{\mathrm{bubble}}+L_{\mathrm{slug}}}\Bigr)^{0.3}.
\end{equation}
This correlation is within $10\%$ accuracy for experimental results for the following range of
parameters: bubble velocity $0.4\,\mathrm{m/s}<U_{\mathrm{bubble}}<2\,\mathrm{m/s}$, bubble length
$1.4<\frac{L_{\mathrm{bubble}}}{d_h}<6.3$, slug length $1<\frac{L_{\mathrm{slug}}}{d_h}<3.2$,
hydrodynamic diameter $d_h = 400
\mathrm{\mu m}$.

\item[II]
For long slug lengths in circular capillaries, where the mass transfer from bubble caps is
predominant, the folowing correlation is derived by
\citet{bercic-mass}:
\begin{equation}
k_L a = \frac{0.111
(U_{\mathrm{gas}}+U_{\mathrm{liq}})^{1.19}}{\bigl((1-\varepsilon_{\mathrm{gas}})(L_{\mathrm{bubble}}
+L_ {\mathrm{slug}} )\bigr)^{0.57} },
\end{equation}
where $\varepsilon_{\mathrm{gas}}=\frac{U_{\mathrm{gas}}}{U_{\mathrm{bubble}}}$ is the gas holdup,
$U_{\mathrm{gas}}$ is the superficial gas velocity, and $U_{\mathrm{liq}}$ is the superficial
liquid velocity. The superficial liquid velocity is calculated with the steady state condition in
the middle of the slug as:
\begin{equation}
U_{\mathrm{liq}}=\frac{\oint U \mathrm{d}A}{A}.
\end{equation}
The superficial liquid velocity can be calculated through the calculated bubble volume and gas
holdup. For the two-dimensional case it is represented as:
\begin{equation}
\begin{aligned}
&\varepsilon_{\mathrm{gas}}=\frac{\text{bubble area}}{15 H^2}\\
&U_{\mathrm{gas}}=\varepsilon_{\mathrm{gas}} U_{\mathrm{bubble}}.
\end{aligned}
\end{equation}

Since experiments done by \citeauthor{bercic-mass} were only for one diffusivity coefficient
$D=2\times 10^{-9}\,\mathrm{m^2/s}$, the effective equation should be rescaled as:
\begin{equation}
k_L a = \sqrt{\frac{D}{2\times 10^{-9}\,\mathrm{m^2/s}}}\frac{0.111
(U_{\mathrm{gas}}+U_{\mathrm{liq}})^{1.19}}{\bigl((1-\varepsilon_{\mathrm{gas}})(L_{\mathrm{bubble}}
+L_ {\mathrm{slug}} )\bigr)^{0.57} },
\end{equation}

\item[III]
\citet{vanbaten-circular} performed numerical simulations for circular capillaries by assuming that
there is a separate contribution of the liquid film and the bubble caps to mass transfer. Their
correlation is as:
\begin{equation}
k_L a = \frac{2}{\sqrt{\pi}}\sqrt{\frac{D U_{\mathrm{bubble}}}{(L_{\mathrm{bubble}}-d_h)}}
\frac{4(L_{\mathrm{bubble}}-d_h)}{d_h(L_{\mathrm{bubble}}+L_{\mathrm{slug}})}+2\frac{\sqrt{2}}{\pi}
 \sqrt{\frac{D U_{\mathrm{bubble}}}{d_h}} \frac{4}{(L_{\mathrm{bubble}}+L_{\mathrm{slug}})}.
\end{equation}
This correlation is valid for short contact times, when the mass transfer is important from bubble
caps as well as from the film. Short contact times are defined through Fourier number as:
\begin{equation}
Fo=\frac{D t_{\mathrm{film}}}{\delta^2},
\end{equation}
where $t_{\mathrm{film}}=\frac{L_{\mathrm{film}}}{U_{\mathrm{bubble}}}$ is the time exposure of the
bubble while it propagates through distance $L_{\mathrm{film}}$. The Fourier number is assumed to
be small if $Fo<0.1$. If $Fo>1$ then the liquid film is saturated and the mass transfer occurs from
the bubble caps.

\end{description}

Overall work goal is to compare numerical simulations with given correlations and to establish a
new one.

\section{Mass transfer}
The mass transfer is characterized by non-dimensional numbers as Schmidt number (viscous diffusion
rate over mass diffusion rate), the Peclet number (convection over diffusion) and the Sherwood
number (convection mass transfer over diffusion mass transfer):
\begin{equation}
\begin{aligned}
&Pe=\frac{U L}{D}\\
&Sh=\frac{K L}{D}\\
&Sc=\frac{\nu}{D}\\
\end{aligned}
\end{equation}

@@

In what follows we will perform the matching of obtained hydrodynamics results
\cite{kuzmin-binary2d} with physical ones from experiment of \citet{bercic-mass}. The diffusion
coefficient of the methane under standard conditions \cite{methane-properties} is
$1.84\times 10^{-5} \mathrm{cm}^2/\mathrm{s}$. The diameters of capillaries in the experiment were
as $1.5$, $2.5$ and
$3.1\,\mathrm{mm}$. For simplicity reasons we will choose the diameter of capillary as $1.5\,
\mathrm{mm}$. 
The non-dimensional parameters obtained from the hydrodynamics lattice Boltzmann simulations are
as:
\begin{equation}
\begin{aligned}
&Ca=\frac{U \mu_{\mathrm{liq}}}{\gamma}\\
&Re=\frac{\rho_{\mathrm{liq}} U L}{\mu_{\mathrm{liq}}}.
\end{aligned}
\end{equation}
Given some physical characteristics as the liquid density and the channel length,
one can obtain the physical velocity obtained from non-dimensional numbers:
\begin{equation}
\begin{aligned}
Ca\cdot Re= \frac{\rho_{\mathrm{liq}} U^2 L}{\gamma}\\
U=\sqrt{\frac{Ca\,Re\,\gamma}{\rho_{\mathrm{liq}}L}}
\end{aligned}
\end{equation}

Assuming that the surface tension of water is $0.0728\,\mathrm{N/m}$ (methane exhibits the same
trend \cite{schmidt-methane}), the density is       
$1000\,\mathrm{kg/m^3}$, the hydraulic diameter of the channel to be $1.5\,\mathrm{mm}$ one can
calculate the physical velocity of the bubble.
Table \ref{table:twod:simulations} represents lattice Boltzmann simulations results and their
physical equivalents for the grid $202\times 3002$. One can see
that velocity is quite reasonable
as \citet{bercic-mass} declared that experiments were conducted for velocities in the ranges of
$0.01$ to $0.4\,\mathrm{m}/\mathrm{s}$. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$    &$Re$     &$U_{LB}$ &$\delta$&$\varepsilon_{\mathrm{gas}}$
&$U_{\mathrm{phys}}, \mathrm{m/s}$\\
\hline
$0.026$ &$0.449$  &$0.0014$ &$0.040$ &$0.306$ &$0.023$   \\ 
$0.047$ &$0.820$  &$0.0027$ &$0.058$ &$0.293$ &$0.043$   \\ 
$0.080$ &$1.378$  &$0.0045$ &$0.085$ &$0.280$ &$0.073$ \\
$0.065$ &$1.126$  &$0.0037$ &$0.076$ &$0.266$ &$0.059$      \\
$0.222$ &$3.807$  &$0.0125$ &$0.122$ &$0.253$ &$0.202$  \\
$0.479$ &$8.222$  &$0.0271$ &$0.151$ &$0.249$ &$0.437$  \\
$0.736$ &$12.617$ &$0.0416$ &$0.164$ &$0.236$ &$0.671$  \\ 
$0.989$ &$16.960$ &$0.0559$ &$0.172$ &$0.230$ &$0.902$  \\
\hline
\end{tabularx}
\caption{Two-dimensional simulations for different
capillary numbers. \label{table:twod:simulations}}
\end{table}

\section{Lattice Boltzmann implementation}
For the lattice Boltzmann implementation we took $\omega=1.99$ which results in the following
lattice Boltzmann diffusivity parameter:
\begin{equation}
\begin{aligned}
&D=\frac{1}{3}\Bigl(\frac{1}{\omega}-\frac{1}{2}\Bigr)\\
&D=0.0008375
\end{aligned}
\end{equation}
In terms of the Schmidt number (taking the lattice Boltzmann liquid viscosity in former simulations
as $\frac{2}{3}$) that
can result in:
\begin{equation}
\begin{aligned}
&Sc=\frac{\nu_{\mathrm{liq}}}{D}\\
&Sc=796\\
\end{aligned}
\end{equation}
Taking the water dynamic viscosity as $8.9 \times 10^{-4}\,\mathrm{Pa\cdot s}$ or kinematic
viscosity as $8.9 \times
10^{-7} \,\mathrm{m^2/s}$ and the Schmidt number as $796$ results in the following physical
diffusion coefficient:
\begin{equation}
D=1.118\times 10^{-5}\,\mathrm{cm^2/s},
\end{equation}
which is less than the diffusivity of the methane $1.84\times 10^{-5}\,\mathrm{cm^2/s}$ but is
physical. After calculation of physical parameters one can calculate correlations indicated in
Section \ref{sec:correlations} for mass transfer. Table \ref{table:lengths} presents extraction
from the lattice Boltzmann simulation necessary to calculate mass transfer calculations.
For the two-dimensional case parameter  $a$ is defined as:
\begin{equation}
a = \frac{\text{bubble perimeter}}{\text{cell area}}.
\end{equation}
For the height of the channel to be $H=1.5 \times 10^{-3}\,\mathrm{m}$, parameters $a$ are
presented in Table \ref{table:lengths}. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$&Bubble length& Slug Length&
$U_{\mathrm{gas}},\,\mathrm{m/s}$&$U_{\mathrm{liq}},\,\mathrm{m/s}$&$a,\,\mathrm{m^{-1}}$\\
\hline
$0.026$&$5.225$&$9.775$ &$0.0070$ &$0.0209$&$510.77$\\
$0.047$&$5.22$ &$9.78$  &$0.0126$ &$0.0379$&$508.62$\\
$0.080$&$5.215$&$9.785$ &$0.0204$ &$0.0615$&$506.53$\\
$0.065$&$4.965$&$10.035$&$0.0157$ &$0.0504$&$482.84$\\
$0.222$&$5.25$ &$9.75$  &$0.0511$ &$0.1560$&$505.06$\\
$0.479$&$5.565$&$9.435$ &$0.1089$ &$0.3155$&$530.37$\\
$0.736$&$5.51$ &$9.49$  &$0.1584$ &$0.4677$&$525.38$\\
$0.989$&$5.505$&$9.495$ &$0.2074$ &$0.6190$&$526.80$\\
\hline
\end{tabularx}
\caption{The nondimensional (scaled to the height of the channel) lengths for the same
lattice Boltzmann simulations as in Table \ref{table:twod:simulations}. As well superficial
gas and liquid velocities are presented. \label{table:lengths}} 
\end{table}

The literature correlations of mass transfer dependence on the bubble velocities are presented in
Fig. \ref{fig:mass:transfer:theoretical}. Note, that the correlations presented in
Fig. \ref{fig:mass:transfer:theoretical} are used primarily for circular capillaries
(\citeauthor{vanbaten-circular,bercic-mass}) and for square capillaries (\citeauthor{yue-mass}). We
adapt these results for flow between parallel plates by using hydraulic
diameter \cite{bercic-mass} which can be calculated as:
\begin{equation}
d_h = \lim_{W\to\infty}\frac{4A}{P}=\lim_{W\to\infty}\frac{4 W\,H}{2(W+H)}=2 H.
\end{equation}
For comparison reasons we as well present the correlation when the hydraulic diameter is taken as
$d_h=H$, see Fig. \ref{fig:mass:transfer:theoretical}.

\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/theoretical_correlations.eps}\\
\includegraphics[width=\textwidth]{Figures/theoretical_correlations_without_hydraulic.eps}\\
\caption{The literature correlations by various authors with hydraulic diameter taken as $2 H$
(top) and $H$ (bottom). Correlations deviate a lot among each
other.\label{fig:mass:transfer:theoretical}}
\end{figure}

The volumetric mass transfer coefficient is calculated according to the work of
\citet{vanbaten-circular}. The concentration flux is calculated as the difference between overall
average concentration taken in the whole domain ($C_{overall}=\int_{V} C \mathrm{d}V /V$)
at time
$t_1$ and time $t_2$ divided on the time difference $t_2-t_1$ {\color{red} (In this particular
moment I don't know whether to take the volume of the whole domain or
just the liquid phase. I will go for whole domain.)}. Then the volumetric mass transfer
coefficient is calculated as \cite{vanbaten-circular}:
\begin{equation}
\label{main:simulation:equation}
k_L a=\frac{\text{Flux}}{C_s-C_{outlet}} \frac{\text{bubble surface area}}{\text{unit cell volume}},
\end{equation}
where $C_{outlet}=\int{C U_{outlet}\mathrm{d}A}/\int{U_{outlet}\mathrm{d}A}$. 
{\color{red} Do not understand why we need this ratio bubble surface area/unit cell volume. In this
case the dimensional units do not coincide. In the calculations I will avoid it until further
clearance.} The overall concentration in the lattice Boltzmann system is prescribed as:
\begin{equation}
C_{overall}=\frac{\sum{c_i}}{N_x \, (N_y-2)}.
\end{equation}

To properly do
non-dimensioanalization one needs to insure the time conversion factor when the flux is calculated
inside the lattice Botlzmann system. The time conversion can be easily obtained:
\begin{equation}
\begin{aligned}
&U_{LB}=U_{phys}\frac{\Delta t}{\Delta x}\\
&\Delta t=\frac{U_{LB}}{U_{phys}}\Delta x\\
&\Delta t=\frac{1.5\,\mathrm{mm}}{200} \frac{0.0559}{0.902\,\mathrm{m/s}}=4.64\times
10^{-7}\,\mathrm{s} 
\end{aligned}
\end{equation}

The bubble surface length is the perimeter of the bubble in two-dimensional space and it is
calculated with the extraction of the bubble contour and representing it with the polygon. After
that the calculated perimeter is the effective summation of line lengths and can be represented as:
\begin{equation}
P=\sum_i{\sqrt{(x2_i-x1_i)^2+(y2_i-y1_i)^2}},
\end{equation}
where the index $i$ represents the polyline consisting of two points $(x1,y1)$ and $(x2,y2)$.
%The corresponding volumetric mass transfer coefficients depending on the velocity are indicated in
%Table \ref{table:volumetric:coefficients}.

\section{Simulation results}
\subsection{Steady-state}
Using Eq.\ref{main:simulation:equation} we calculated the mass transfer coefficient. One can see
the plots for the mass transfer coefficient versus time in Fig. \ref{fig:steady:state}. We state as
a steady-state condition the average of concentration values between $500000$ to $1000000$ time
steps.{\color{red} We need to put more time steps. As well it is probably connected with Fourier
number.} The average calculated volumetric mass transfer coefficients values are presented in Table
\ref{table:volumetric:coefficients}. For each of the measurements we as well provided the Fourier
number (for the flow between plates is as follows):
\begin{equation}
Fo=\frac{D t_{\mathrm{film}}}{\delta^2}=\frac{D
(L_{\mathrm{bubble}}-H)}{U_{\mathrm{bubble}}\delta^2}.
\end{equation}
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/steady_state.eps}
\caption{The mass transfer coefficient vs. time iterations. One can see that the pseudo steady
state can be reached after $500000$. The mass transfer coefficient is taken as the average of
concentration values between $500000$ to $1000000$ time steps.\label{fig:steady:state}}
\end{figure}
\begin{table}[htb!]
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
$Ca$&$Fo$&$k_L a,\,\mathrm{s}^{-1}$\\
\hline
$0.026$&$0.08493$&$0.2737$\\
$0.047$&$0.02160$&$0.3664$\\
$0.080$&$0.00588$&$0.4959$\\
$0.065$&$0.00862$&$0.4734$\\
$0.222$&$0.00103$&$0.8505$\\
$0.479$&$0.00034$&$1.1541$\\
$0.736$&$0.00018$&$1.1801$\\
$0.989$&$0.00012$&$1.2215$\\
\hline
\end{tabularx}
\caption{Simulation results for different capillary numbers. \label{table:volumetric:coefficients}}
\end{table}

Comparison between correlations and simulations one can find in Fig.
\ref{fig:comparison:correlations}.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/comparison_correlations.eps}
\caption{One can see the comparison between correlations and simulations. {\color{red} Last two
points can be eliminated due to improper imposement of the concave structure of the bubble. The
large deviation of the results can be probably explained by the flow between plates in comparison
with the flow in circular and square microchannels.} \label{fig:comparison:correlations}}
\end{figure}
